---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fishEVA

<!-- badges: start -->
<!-- badges: end -->

The goal of fishEVA is to use Extreme Value Analysis (EVA) to estimate the maximum body length of fishes. The packages uses two approaches, the first is from Extreme Value Theory (EVT), which shows that the maxima of a set of samples follows a specific distribution - the Generalised Extreme Value (GEV) distribution. The second approach uses knowledge on the underlying body size distribution to estimate the likely parameters of the underlying distribution that would give rise to the observed sample maxima. The two approaches are implemented using a Bayesian Framework. 

## Installation

You can install the development version of fishEVA from [GitHub](https://github.com/) with:

```{r}
# install.packages("pak")
pak::pak("FreddieJH/fishEVA")
```

## Example

This is a basic example which shows you how to fit an EVT model to a set of maxima values:

```{r example}
library(fishEVA)
## basic example code
length_maxima <- c(40, 41, 35, 42, 31) #cm

# model_fit <- fit_mod(length_maxima, model_type = "evt")

```

